---
title: "QMM-ASSIGNMENT3"
author: "Sandhya"
date: "2024-10-04"
output: html_document
---
Question:

Heart Start produces automated external defibrillators (AEDs) in each of two different plants (A
and B). The unit production costs and monthly production capacity of the two plants are
indicated in the table below. The AEDs are sold through three wholesalers. The shipping cost
from each plant to the warehouse of each wholesaler along with the monthly demand from each
wholesaler are also indicated in the table. How many AEDs should be produced in each plant,
and how should they be distributed to each of the three wholesaler warehouses so as to minimize
the combined cost of production and shipping?
Unit Shipping Cost Unit Monthly
Warehouse 1 Warehouse 2 Warehouse 3 Production
Cost
Production
Capacity
Plant A $22 $14 $30 $600 100
Plant B $16 $20 $24 $625 120
Monthly
Demand
80 60 70
Formulate and solve this transportation problem.
Requirements


```{r}
# Install and load lpSolve
if (!requireNamespace("lpSolve", quietly = TRUE)) {
  install.packages("lpSolve")
}
library(lpSolve)
```


```{r}
# Creating a dataframe for the transportation problem
datafr <- data.frame(Warehouse_1 = c("$22", "$16", 80),
                     Warehouse_2 = c("$14", "$20", 60),
                     Warehouse_3 = c("$30", "$24", 70),
                     Unit_production_cost = c("$600", "$625", " "),
                     Monthly_production_capacity = c(100, 120, " "))
rownames(datafr) <- c("Plant A", "Plant B", "Monthly Demand")

# Generating a formatted table
library(knitr)
#install.packages("kableExtra")
library(kableExtra)

datafr %>%
  kable() %>%
  kable_classic() %>%
  column_spec(1, border_right = TRUE) %>%
  column_spec(4, border_right = TRUE) %>%
  row_spec(2, extra_css = "border-bottom:dotted") %>%
  add_header_above(header = c(" " = 1, "Unit Shipping Cost" = 3, " " = 2))

```


  Problem Setup:
  Plants: A and B
  Warehouses: 1, 2, 3

	Production Costs:
	Plant A: $600 per unit
	Plant B: $625 per unit
	
	
	Shipping Costs:
	Plant A to Warehouse 1: $22, Warehouse 2: $14, Warehouse 3: $30
	Plant B to Warehouse 1: $16, Warehouse 2: $20, Warehouse 3: $24
	
	
	Monthly Production Capacity:
	Plant A: 100 units
	Plant B: 120 units
	
	
	Monthly Demand:
  Warehouse 1: 80 units
  Warehouse 2: 60 units 
  Warehouse 3: 70 units

Decision Variables:


Let X11, X12, X13 represent the number of AEDs shipped from plant A to warehouse 1, 2, 3 respectively, and X21, X22, X23 represent same from plant B.

Objective function-

Here We are aiming to minimize the total cost of production and shipping,which we can write as:

Min Z = (600+22)x11 + (600+14)X12 + (600+30)x13 + (625+16)x21 + (625+20)x22 + (625+24)x23

This simplies to :

MIn Z =622x11 + 614x12+ 630x13 + 641x21 + 645x22 + 649x23


Constraints-

$$1) X11 + X12 + X13 \le 100 (Plant A capacity)$$

$$2) X21 + X22 + X23 \le 120 ( plant B capacity)$$

$$3) X11 + X21 \ge 80 (warehouse 1 demand)$$

$$4) X12 + X22 \ge 60 (warehouse 2 demand)$$

$$5) X13 + X23 \ge 70 (warehouse 3 demand)$$

Non- Negativity-

$$All\ Xij \ge 0.$$





```{r}
# Defining the costs associated with production and shipping.
# Decision variables order: 
# [A to Warehouse 1, A to Warehouse 2, A to Warehouse 3, 
#  B to Warehouse 1, B to Warehouse 2, B to Warehouse 3]
Shipping_costs <- c(600 + 22, 600 + 14, 600 + 30, 
           625 + 16, 625 + 20, 625 + 24)

```



```{r}
# Setting up the matrix that will include all the constraints for the problem.
# Each row corresponds to a constraint: 
# 1) Plant A capacity, 2) Plant B capacity
# 3) Warehouse 1 demand, 4) Warehouse 2 demand, 5) Warehouse 3 demand
constraint_matrix <- matrix(c(1, 1, 1, 0, 0, 0,   # Capacity of Plant A
                              0, 0, 0, 1, 1, 1,   # Capacity of plant B
                              1, 0, 0, 1, 0, 0,   # Demand at the warehouse 1
                              0, 1, 0, 0, 1, 0,   # Demand at the warehouse 2
                              0, 0, 1, 0, 0, 1),  # Demand at the warehouse 3
                            nrow = 5, byrow = TRUE)

```



```{r}

# Defining the Right-hand side values for the constraints
constraints_rhs <- c(100, 120, 80, 60, 70)  # Values of Capacity and Demand

```



```{r}
# Specifying the directions for the constraints
# "less than or equal to" for the capacity of plant, "equal to" for the warehouse demands
directions <- c("<=", "<=", "=", "=", "=")
```


```{r}
# Solving the transportation problem using lp() function
solution <- lp("min", Shipping_costs, constraint_matrix, directions, constraints_rhs)
solution

```



```{r}
# Displaying the results
if (solution$status == 0) 
  # If an optimal solution is found then print the results
  print("Optimal solution found:")
  solution$solution_matrix <- matrix(solution$solution, nrow = 2, byrow = TRUE)
  rownames(solution$solution_matrix) <- c("Plant A", "Plant B")
  colnames(solution$solution_matrix) <- c("Warehouse 1", "Warehouse 2", "Warehouse 3")
  print(solution$solution_matrix)

```





 # Calculating the total number of AEDs to be produced by Plant A
 
```{r}
total_A <- sum(solution$solution_matrix[1, ])
  print(paste("Number of AEDs to be produced by Plant A: ", total_A))

```
 

```{r}
 
  # Calculating the total number of AEDs to be produced by Plant B
  total_B <- sum(solution$solution_matrix[2, ])
  print(paste("Number of AEDs to be produced by Plant B: ", total_B))
  
```
  
  
```{r}

 # Display the minimized total cost
  print(paste("Minimized total cost: $", solution$objval))
```
  
 
  
  
```{r}
# Total cost
print(paste(solution$objval))
```
Conclusion:


The transportation problem for Heart Start has been optimized using linear programming, enabling efficient distribution of automated external defibrillators (AEDs) from two plants to three wholesalers. Plant A will produce 100 units, while Plant B will produce 110 units, meeting all monthly demands within capacity limits. The total minimized cost for production and shipping is $132,790. This optimization streamlines logistics and enhances resource utilization. Ultimately, it allows Heart Start to reduce operational costs and improve customer service, strengthening its market competitiveness.



